---
title: "Assignment 2"
author: "Marco Palombo"
date: "10/25/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
```

# Exercise 1
In winter, a home in Phoenix, AZ, requires 62kWh of heat per day to remain at a constant indoor
temperature of 20°C.

a. The average daily solar radiation, i.e. the average over a 24-hour period, in Phoenix in Winter
is 270W/m2. What is the minimum collector surface needed for an all-solar heating system that
has an overall efficiency of 20%?

```{r}
# part 1b
surface_area <-  62/24*0.2/0.27
```


$$ A~=~62~kWh*\frac{1}{24~h}*0.2*\frac{m^2}{0.27~kW}~=~1.9~m^2 $$

b. Water is used as the heat storage medium. How much water is required if it is heated from 20°C
to 60°C every day

NOTE TO SELF THIS RESULT SEEMS HUGE CHECK WITH OTHERS

```{r}
total_energy <-  62*3.6e6/4184/40
```
$$ 62~kWh*\frac{3.6e^6~J}{1~kwh}*\frac{kg~K}{4184~J}*\frac{1}{40~K}~=~1.3333~kg $$

# Exercise 2

a) An automobile weighing 4409 pounds decelerates from 55mph to 0 m/s using mechanical
brakes. How much Joules of heat is generated by the brakes?

```{r}

```


First convert 55 mph to m/s and 4409 pounds to kg
$$ 55~mph *\frac{0.44~\frac{m}{s}}{mph}~=~24.2~\frac{m}{s}$$
$$ 4409~lbs *\frac{0.453~kg}{1~lbs}~=~1997~kg$$
```{r}
kinetic_energy <- 1997*24.2^2/2
```


Complete conversion of kinetic energy to heat energy in brakes
$$ Energy~=~\frac{m*v^2}{2}~=~\frac{(1997~kg)*(24.2 \frac{m}{s})^2}{2}~=~ 585~kJ$$

b) Each of the four brake drums has a mass 9kg of iron (specific heat: 0.11 kcal/kg∙°C). By how much does the temperature of the brake drums rise (assuming that each drum absorbs the same amount of energy, and no heat is transferred to the air or other vehicle parts)?

Convert specfic heat into $\frac{J}{Kg~K} $
$$ 0.11 \frac{Kcal}{kg~K}*\frac{4.184~kJ}{1~kcal}~=~ 0.46~\frac{kJ}{kg~K}$$

Divide total energy by 4 to account for distribution across 4 brake systems then divide by thermal mass:
```{r}
temp_change <- 584/4/0.46/9

```
$$ 584~kJ*\frac{1}{4}*\frac{kg~K}{0.46~kJ}*\frac{1}{9~kg}~=~35~K$$

# Exercise 3
A cylinder contains 1 kg of carbon dioxide, which has a temperature of T = 293 K (20 ºC) and is at normal atmospheric pressure p = 1 atm or 101,325 N/m2. Assume that the carbon dioxide behaves like an ideal gas. The molar weight of CO2 is 44 g/mol.

a) What is the volume V of the carbon dioxide at 293 K and 1 atm?

First determine quantity of mols:
$$ 1~kg*\frac{mol}{.044~kg}~=~ 2.27~mol$$

Use ideal gas law equation:
$$ PV=nRT $$

```{r}
Volume = 2.27*8.314*293/101325
```

Solved for Volume:
$$ V=\frac{nRT}{P}~=~\frac{(2.27~mol)(8.314\frac{J}{mol~K})(293~K)}{101,325~\frac{N}{m^2}}~=~0.545~m^3 $$

b) To what temperature does the carbon dioxide heat up if it is adiabatically compressed to a tenth of its original volume?

Apply adiabatic equation:
$$ C_v \frac{1}{2} $$

Solve for $T_2$:
$$ T_2~=~\frac{R}{C_v}*\ln $$
